### 2014-02-10 thesis/src/chapters/chapter-adaptivity.tex
Thesis: adaptivity section has been rewritten

\section{Self-adaptive system}

\subsection{Introduction}
What makes a concept of an adaptivity enticing is crucial idea behind guaranteeing Quality-of-Serivce: automatization. Generally, ensuring and enforcing given quality can be seen as continuous monitoring and reacting to events when necessary. Such idea can be expressed as IT process shown in figure \ref{fig:it-process}.

@@

While system administrator can be directly responsible for manually performing above-mentioned steps, it is much more favourable to automatize this process leading to a system self-management or an adaptable system. Not only does it lead to efficiency of an IT process but also to its effectiveness \cite{IBM06}. It is possible through:
\begin{itemize}
  \item \emph{rapid process initiation} - components auto-initiates actions based on information derived from a system
  \item \emph{reduced time and skill requirements} - automatization of IT processes makes them easier and less troublesome what is especially important for skill-intensive, error-prone and long lasting tasks
\end{itemize}

In a context of delivering a computing platform, adaptivity adds auto-scaling features to a solution offered by Platform-as-a-Service provider. Self-managing is possible through usage of an Elasticity Controller which gathers probes from resource such as application container and uses that knowledge to execute appropriate action on that resource, indirectly modifying consecutive probes \cite{VaRoBu11}. This concept illustrates diagram \ref{fig:elasticity-controller}.

\begin{figure}[!ht]
  \begin{center}
    \includegraphics{chapter-adaptivity/elasticity-controller}
  \end{center}
  \caption{Elasticity controller}
  \label{fig:elasticity-controller}
\end{figure}

The remaining of this chapter describes crucial elements that compose elasticity controller: policies, data analysis, triggered actions and presents a comparison of cloud providers.

\subsection{Policies}
While auto-scaling is offered by a vast amount of cloud providers (e.g AWS, OpenShift, OpenNebula) it often lacks a sophisticated mechanisms allowing for specific scaling policies, being limited to only one predefined rule as it is in case of OpenShift for example.

Policy denotes a condition which, when satisfied, triggers an action that is supposed to harness cloud instance in a way that future evaluations of condition will be unsuccessful. Typically condition itself is accompanied by a minimal and maximal number of node instances, allowing for ensuring minimal QoS and controlling maximal costs. Currently, industry leaders supports two main kind of policies \cite{AmazonAutoScaling}:
\begin{itemize}
 \item \textit{expression based} - allows to define how you to scale application in response to changing conditions, which include factors such as memory, CPU usage, cost or some indirect, calculated metrics
 \item \textit{scheduled} - allows to scale an application in response to predictable load changes. For example, traffic increases during the weekends and decreases on working days. Hence, that predictable traffic patterns is used to scale application based on current time.
\end{itemize}

Technically, policies are expressed in some human-readable format such as JSON, XML as it is in case of AWS EC2 or custom expression used for example by Carina. Appendix \ref{app:scaling-policies} presents example configuration used by AWS E2 Auto-Scaling.

\subsection{Data analysis}
Having policies defined, their conditions are evaluated against data acquired from sensors. In a simplest case this evaluation can be based on a Threshold Model \cite{LiWoZh05}, which defines a valid range. In cases when given metric violates that condition (i.e. value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted - figure \ref{fig:threshold-model} illustrates that idea.

@@

While trivial in its form, cases of AWS, OpenShift, Carina, OneCloud proves it is useful in a real-world scenarios.

\subsection{Triggered actions}
Actions that are being triggered by a elasticity controller are focused on application scaling. Previous chapter described that problem in detail.

>>>>

\section{Self-manageable system}

\subsection{Introduction}
Guaranteeing Quality-of-Service, a critical concept in this paper, can be seen as continuous monitoring and reacting to undesirable conditions when necessary. In fact, this process is as an IT procedure and traditionally belongs to the duties of a system administrator. Figure \ref{fig:it-process} illustrates exemplary IT procedure.

@@

While system administrator can be directly responsible for manually performing steps depicted in figure \ref{fig:it-process}, it is much more favourable to automatize this process leading to a self-manageable system. Not only does it lead to efficiency of an IT process but also to its effectiveness \cite{IBM06} achieved through:
\begin{itemize}
  \item \emph{rapid process initiation} - components auto-initiates actions based on information derived from a system
  \item \emph{reduced time and skill requirements} - automatization of IT processes makes them easier and less troublesome what is especially important for skill-intensive, error-prone and long lasting tasks
\end{itemize}

As  \cite{murch2004autonomic} explains, there are four levels of a system manageability maturity. The higher the level, the better it manages itself leading to a adaptive or autonomic system at best:
\begin{itemize}
 \item \emph{Basic} - IT professionals manage each system element. Activities related to configuration, optimisation, healing, and protection of IT infrastructure components are performed manually.
 \item \emph{Managed} - System management technologies can be used to collect information from different systems and analyze this information. In this scenario, IT professionals do most of the analysis, but it is a starting point for the automation of IT tasks.
 \item \emph{Predictive} - Individual components monitor themselves, analyze changes and offer advice. Therefore, dependency on people is reduced and decision-making is improved.
 \item \emph{Adaptive} - IT components can, individually or group-wise, monitor, analyze operations and offer advice with minimal human intervention.
 \item \emph{Autonomic} - System operations are managed by business policies established by the administrator. In fact, business policy drives overall IT management, while at the adaptive level; there is an interaction between humans and the system.
\end{itemize}

In heart of the every self-adaptive systems, according to \cite{brun2009engineering}, lies a feedback loop, concept that originates from a control engineering. As paper's authors advocate:
\begin{quote}
Feedback loops provide the generic mechanism for self-adaptation. Positive feedback occurs when an initial change in a system is reinforced, which leads toward an amplification of the change. In contrast, negative feedback triggers a response that counteracts a perturbation. 
\end{quote}

What control loop is made of is collection of data, its analysis, making a decisions and enforcing them in the environment. Figure \ref{fig:feedback-loop} depicts a generic feedback loop, giving context associated with each phase.

\begin{figure}[!ht]
  \begin{center}
    \includegraphics[width=\textwidth]{chapter-adaptivity/feedback-loop}
  \end{center}
  \caption{Feedback loop \cite{brun2009engineering}}
  \label{fig:feedback-loop}
\end{figure}

Successive sections details aforementioned phases and gives an examples of systems leveraging feedback loops.

\subsection{Collecting data}
During this phase, relevant data is collected from sensors. After that, data can be aggregated, filtered or stored to provide a reference point during future feedback loop evaluation. During this phase, there are a few concerns that should be investigated \cite{brun2009engineering}:
\begin{itemize}
 \item sample rate
 \item probe reliability
 \item data format
\end{itemize}

Almost any cloud platform providers incorporates into its solution monitoring component, responsible for data collection. Mechanisms used vary from providers to provider and may involve mixture of protocols and standards. For example, Carina \cite{Carina} uses OpenNebula IM and VMM drivers relaying on a plaintext data, transferred through a SSH connection.

\subsection{Analysis}
Analysis part applies mathematical models to reason about data gathered during previous phase. It can leverage variety of approaches and be driven by a customer-defined policies.

As \cite{brun2009engineering} specifies, it is important to take into account following concerns during analysis phase:
\begin{itemize}
 \item applying historical data and existing patterns to current situation
 \item archiving data
 \item applying adequate model to current situation
 \item model's stability
\end{itemize}

Currently, service providers employs only simplified models such as threshold model \cite{LiWoZh05}, that defines a valid range for a given metric. In cases when it is violated (i.e. value is either smaller than minimal or bigger than maximal acceptable) corresponding resource is properly adjusted - figure \ref{fig:threshold-model} illustrates that idea.

@@

Threshold model, while trivial in its form, has been proved useful in a real-world scenarios as it is in case of AWS, OpenShift, Carina or OneCloud.

As it was already mentioned, models are often controlled by a policies that denotes a condition which, when satisfied, triggers an action, harnessing environment's disturbance. Currently, industry leaders supports \cite{AmazonAutoScaling} two kind of policies:
\begin{itemize}
 \item \textit{expression based} - allows to define how to scale application in response to changing conditions, which include factors such as memory, CPU usage, cost or some indirect, calculated metrics
 \item \textit{scheduled} - allows to scale an application in response to predictable load changes. For example, traffic increases during the weekends and decreases on working days. Hence, that predictable traffic patterns is used to scale application based on current time.
\end{itemize}

Technically, policies are expressed in a human-readable format such as JSON, XML as it is in case of AWS EC2 or custom expression used for example by Carina. Appendix \ref{app:scaling-policies} presents XML format adopted by AWS E2 Auto-Scaling.

\subsection{Decision arrangement}
Before implementing changes in an environment, decision what resources have to be tuned in order to reach a desirable state is made. Decision should be made on the basis of:
\begin{itemize}
 \item risk analysis - analysing how change effects whole environment
 \item prioritising identified problems and resources
\end{itemize}

In a case of platform-as-a-service providers, none of them explicitly specifies how a change plan is composed.

\subsection{Actions}
Finally, a change plan is implemented in a system. Actions can cover a variety of scaling and resource tuning operations. Next chapter investigates problem of scaling applications in detail.

### 2014-02-10 thesis/src/chapters/chapter-design.tex
Thesis: adaptivity section has been rewritten

\section{Summary}
Cloud-SAP fills the gap in existing platforms, providing a standardised approach to scaling applications that uses a fine-grained actions, operating on different platform levels: starting from execution environment to end on a cloud federation. It is based on a concept of an autonomic system, providing methodical yet flexible foundation for designed system. Cloud-SAP leverages core benefits of such system: promise of a self-configuring, self-optimising, self-healing, self-protecting system driven by a customer-defined policies.

>>>>

\section{Summary}
Cloud-SAP fills the gap in existing platforms, providing a standardised approach to scaling applications that uses a fine-grained actions, operating on different platform levels: starting from execution environment to end on a cloud federation. It is based on a concept of an autonomic system, providing methodical yet flexible foundation for designed system. What is more, Cloud-SAP leverages core benefits of such system: promise of a self-configuring, self-optimising, self-healing, self-protecting system driven by a customer-defined policies and consequently ensuring that customer is served with required Quality-of-Service level with optimised costs.
